### 2024-04-02 Notes/Notes.tex
Initial commit after a bit of work

\documentclass[12pt]{article}
\usepackage{}
\title{A480 Notes}
\author{Brayden Leicester}
\begin{document}
\maketitle

\section{Things to Check}
The named asteroids in \cite{Harris1994} (most notably 4179 Toutatis) should be checked for TESS observations. As these should tumble.

288 Glauke \cite{Harris2015}

\section{Background writing}
%TODO 1I observation paragraph/s

The first interstellar object discovered in late 2017 has come to be called 1I/'Omuamua. \dots

%TODO TESS intro
The Transiting Exoplanet Survey Satellite (TESS) \cite{Ricker2014} is a large area, high cadence, imaging \dots

\section{Method Notes}

\bibliographystyle{apalike}
\bibliography{bibfile.bib}
\end{document}